---
title: "Reproducible Research: Peer Assessment 1"
author: Nwokedi Idika
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(dplyr)
df <- read.csv('activity.csv')
df_clean <- df[complete.cases(df),]
```


## What is mean total number of steps taken per day?

* Histogram of total steps taken per day

```{r}
total_per_day <- group_by(df_clean, date) %>% summarize(total = sum(steps))
hist(total_per_day$total, main = "Histogram of Total Steps Taken \nPer Day", xlab = "Number of Steps Taken in a Day")
```

* The mean number of steps taken per day is given below

```{r}
mean(total_per_day$total)
```

* The median number of steps taken per day is given below 

```{r}
median(total_per_day$total)
```

## What is the average daily activity pattern?

* Time series 
```{r}
avg_interval <- group_by(df_clean, interval) %>% summarize(avg = mean(steps))
plot(avg_interval$interval, avg_interval$avg, type="l", ylab="Avg. Number of Steps", xlab = "5-Minute Intervals", main = "Average Number of Steps Over Time")
```

* The interval with the largest number of steps 
```{r}
arrange(avg_interval, desc(avg))[1,1]
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
