---
title: "Reproducible Research: Peer Assessment 1"
author: Nwokedi Idika
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(dplyr)
df <- read.csv('activity.csv')
df_clean <- df[complete.cases(df),]
```


## What is mean total number of steps taken per day?

* Histogram of total steps taken per day

```{r}
total_per_day <- group_by(df_clean, date) %>% summarize(total = sum(steps))
hist(total_per_day$total, main = "Histogram of Total Steps Taken \nPer Day", xlab = "Number of Steps Taken in a Day")
```

* The mean number of steps taken per day is given below

```{r}
mean(total_per_day$total)
```

* The median number of steps taken per day is given below 

```{r}
median(total_per_day$total)
```

## What is the average daily activity pattern?

* Time series 
```{r}
avg_interval <- group_by(df_clean, interval) %>% summarize(avg = mean(steps))
plot(avg_interval$interval, avg_interval$avg, type="l", ylab="Avg. Number of Steps", xlab = "5-Minute Intervals", main = "Average Number of Steps Over Time")
```

* The interval with the largest number of steps 
```{r}
arrange(avg_interval, desc(avg))[1,1]
```

## Imputing missing values

* The number of missing values in the dataset is given below
```{r}
nrow(df) - nrow(df_clean)
```

* We'll use the average number of steps for an interval to fill in missing values 

```{r}
no_vals <- df[is.na(df$steps),]
for(i in 1:nrow(no_vals)){
  no_vals[i,1] <- avg_interval[avg_interval[,1] == no_vals[i,3],2]
}

df_filled <- rbind(df_clean, no_vals)
```

* Histogram of new dataset
```{r}
total_per_day_filled <- group_by(df_filled,date) %>% summarize(total=sum(steps))
hist(total_per_day_filled$total, main = "Histogram of Total Steps Taken \nPer Day", xlab = "Number of Steps Taken in a Day")
```

* Mean of the dataset

```{r}
mean(total_per_day_filled$total)
```

* Median of the dataset

```{r}
median(total_per_day_filled$total)
```

The mean and median are now the same when previously they differed some. 

## Are there differences in activity patterns between weekdays and weekends?

```{r}
mk_categories <- function(days){
  r <- NULL
  for(day in days){
    if("Saturday" %in% day || "Sunday" %in% day){
      r <- c(r, "Weekend")
    }
    else{
      r <- c(r, "Weekday")
    }
  }
  r
}
w <- weekdays(as.Date(df_filled$date))
w <- mk_categories(w)
df_new <- data.frame(df_filled, tow=as.factor(w))
df_wk <- group_by(df_new, tow, interval) %>% summarize(avg=mean(steps))

library(ggplot2)
qplot(interval, avg, data=df_wk, facets= tow ~ ., geom = "line", ylab = "Avg. Steps Taken", xlab = "5-minute Intervals")
```
